Sperm should evolve to make female meiosis fair.

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
yb minor modifications to two-locus models

Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
	with alternative models of two loci - one a female driver, 
	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
In this model, a female meiotic driver with no sperm-dependency 
	initially reaches drive-viability equilibrium (with two alleles
	$A$ and $B$ are the ancestral non-driver and driver alleles). 
Subsequently, a sperm-acting modifier of female meiotic drive arises at another lous.

We first assume that the modifier is tightly
        linked to the drive locus (effectively creating a third
        allele/haplotype at this locus) and arises on the drive-background. 
Tight linkage offers the best
        chance for a collaboration to evolve between a driver and
       an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}. 
We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
	in Model 5 the modifier arises tightly linked to the non-driving allele, 
	and in Model 6 it is unlinked to the driver.

@@

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. \gc{This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes.} This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.

>>>>

Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
	with alternative models of two loci - one a female driver, 
	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
In this model, a female meiotic driver with no sperm-dependency 
	initially reaches drive-viability equilibrium (with two alleles
	$A$ and $B$ are the ancestral non-driver and driver alleles). 
Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.

We first assume that the modifier is tightly
        linked to the drive locus (effectively creating a third
        allele/haplotype at this locus) and arises on the drive-background. 
Tight linkage offers the best
        chance for a collaboration to evolve between a driver and
       a sperm-acting drive  enhancer, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}. 
We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
	in Model 5 the modifier arises tightly linked to the non-driving allele, 
	and in Model 6 it is unlinked to the driver.

@@

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and Equations \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.
